---
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: F
    theme: journal
    code_folding: hide
params:
  
  report_author: NULL
  report_title: Calendar Item Visualization Report
  daynamica_data: NULL
  user_id_list: NULL
  limit_tod: NULL
  color_codings: NULL
editor_options: 
  chunk_output_type: console
---

---
title: `r params$report_title`
author: `r params$report_author`
date: `r Sys.Date()`

---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, message = F, warning = F}
library(dplyr)
library(ggplot2)

source('../helper_functions/s4_temporal_plot.R')
library(lubridate)
library(ggnewscale)
library(wesanderson)
daynamica_data <- params$daynamica_data
if(is.null(params$color_codings)){


  unique_activity_labels <- 
    daynamica_data$ucalitems_temporal_plot %>%
    tibble() %>%
    filter(type_decoded == "ACTIVITY") %>%
    pull(subtype_decoded) %>% unique()
  unique_trip_labels <- 
    daynamica_data$ucalitems_temporal_plot %>%
    tibble() %>%
    filter(type_decoded == "TRIP") %>%
    pull(subtype_decoded) %>% unique()
  
activity_palette <- c(wes_palette("Zissou1", 1),
                      wes_palette("Zissou1", 5)[4:5],
                      "#679669", # Green
                      "#876796", # Purple
                      "#96676F", # Dark red purple
                      rev(wes_palette("Royal1", 4)[c(4)]),
                      wes_palette("Moonrise3", 2),
                      wes_palette("GrandBudapest2"),
                      wes_palette("GrandBudapest1"),
                      wes_palette("Chevalier1", 4))[1:length(unique_activity_labels)]
  
  
  trip_palette <- c(palette.colors(n = 8, "ggplot2")[-1],
                    palette.colors(n = 8, "alphabet"))[1:length(unique_trip_labels)]
  
  
  color_codes_activity_data <- data.frame(labels = unique_activity_labels,
                                     values = activity_palette)
  
  color_codes_trips_data <- data.frame(labels = unique_trip_labels,
                                  values = trip_palette)
}else{
  color_codes_activity_data <- params$color_codings$color_codes_activity_data
  color_codes_trips_data <- params$color_codings$color_codes_trips_data
}

if(is.null(params$limit_tod)){
  min_time = 0
  max_time = 1
} else{
  min_time <- limit_tod[1]
  max_time <- limit_tod[2]
}
```

# Visualization Notes

These plots show 7-day views of the calendar item data collected via the Daynamica app. Each figure starts on a Sunday. For participants who collected data for longer than one week, there will be multiple plots which can be accessed via separate tabs. 

Colors and Category levels can be modified via dataset manipulations and report parameters. See the master_analysis_file.R for further details. 

# User Calendar Item Visualizations {.tabset}



```{r, results = 'asis', fig.width = 14, fig.height = 10}
pages <- ceiling(length(user_id_list) /10)
user_id_list_short <- sub("@[^@]*$", "", user_id_list)
for(i in 1:pages){
  if(i != pages){
    id_index <- (1 + 10*(i-1)):(10 + 10*(i-1))
  } else{
    id_index <-(1 + 10*(i-1)):length(user_id_list)
  }
  
  cat("\n\n## Participants: ", min(id_index), "-", max(id_index), " {.tabset} \n")
  for(j in id_index){
    cat("\n\n### ", user_id_list_short[j], " {.tabset} \n")
      #print(user_id_list_short[j])
      
      temp_data <- daynamica_data$ucalitems_temporal_plot %>%
        as_tibble() %>%
        filter(user_id == user_id_list[j])
      
      tab <- tibble(start_date = as.Date(min(temp_data$start_date):max(temp_data$start_date), origin = "1970-01-01"),
             day_number = lubridate::wday(start_date),
             check = ifelse(day_number == 1, 1, 0),
             cum_total = cumsum(check))
      
      
      # tab <- temp_data %>%
      #   select(start_date, dow) %>%
      #   mutate(day_number = lubridate::wday(start_date)) %>%
      #   distinct() %>%
      #   arrange(start_date) %>%
      #   mutate(check = ifelse(day_number == 1, 1, 0),
      #          cum_total = cumsum(check))
      
      unique_weeks <- unique(tab$cum_total)
      for(k in 1:length(unique_weeks)){
        current_dates <- tab %>%
          filter(cum_total == unique_weeks[k])
        suppressWarnings(dates_this_week <- as.Date((current_dates$start_date - current_dates$day_number + 1) + 0:6, origin = "1970-01-01"))
        
        #tab_text <- paste0(min(current_dates$start_date), " : ", max(current_dates$start_date))
        tab_text <- paste0(min(dates_this_week), " : ", max(dates_this_week))
        cat("\n\n#### ", tab_text, " \n")
        analysis_data <- temp_data %>%
          filter(start_date %in% dates_this_week)
        if(nrow(analysis_data) == 0){
          print("No data during this date range")
          next
        }
          analysis_data <- analysis_data %>%
            bind_rows(analysis_data[rep(1, 7),] %>% # This code allows all days of the week to be displayed even if we don't have data on those dates.
                      mutate(start_time = "04:00:00",
                             end_time = "04:00:01",
                             start_date = dates_this_week,
                             type_decoded = "ACTIVITY")) %>%
            mutate(end_time = ifelse(end_time == "00:00:00", "24:00:00", end_time))
      p <- suppressMessages(generate_time_of_day_plot(analysis_data,
                           color_codes_activity = color_codes_activity_data,
                           color_codes_trips = color_codes_trips_data,
                           min_time = min_time,
                           max_time = max_time)) + 
        ggtitle(paste0("Calendar Item Visualization: ", user_id_list_short[j]),
                subtitle = paste0("Date Range: ", tab_text))
      print(p)
      }
  }
}

```